---
title: "Limited English Profiency on Income Levels"
execute: 
  echo: false
---

```{r}
#| message: false

library(tidyverse)
library(tidymodels) # For mathematical models (i.e Linear Regression)
library(broom) # includes tidy()
library(marginaleffects) # To compare actual values and predictions
library(scales)
```

```{r}
# Read data
surveys <- read_rds(file = "data/usa_2021_23.rds")
```

```{r}
# Using income wage as a linear function of LEP
# Controlling other variables, such as a person's education and sex

fit_income <- linear_reg() |>
  fit(log_income ~ lep + educ_level + gender, data = surveys)
```

```{r}
# Predictions of data
predictions <- plot_predictions(
  fit_income,
  condition = c("lep", "educ_level", "gender"),
  draw = FALSE
)

# Visualize predictions using ggplot2
# Plot predicted income (back-transformed from log) with all variables

predictions |>
  ggplot(
    aes(
      x = educ_level,
      y = expm1(estimate),  # Subtract 1 to reverse the +1 from log transformation
      fill = gender
    )
  ) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_errorbar(
    aes(ymin = expm1(conf.low), ymax = expm1(conf.high)),  # Also adjust error bars
    position = position_dodge(width = 0.9),
    width = 0.25
  ) +
  facet_grid(
    ~ lep,
    labeller = labeller(
      lep = c(`1` = "Limited English", `0` = "English Proficient")
    )
  ) +
  labs(
    title = "Predicted Income by LEP, Education Level, and Gender",
    x = "Education Level",
    y = "Predicted Income (USD)",
    fill = "Gender"
  ) +
  scale_y_continuous(labels = dollar_format(scale = 1)) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

Language proficiency is one of the key requirements in almost every job posting, and being proficient in a language can open new doors for an individual. Using a sample of data drawn from the American Community Surveys (U.S. Census) of 2021-2023, I seek to examine the effect of Limited English proficiency on a person's income in the USA. I used R to look at the insights into how Limited English Proficiency (LEP) effects a person's wage when controling education level and gender. I modeled a person's income, as a linear function of her English profiency, education level, and gender. Across all education levels, males earn 20–50% more than females, with the largest gender and language-based income gaps appearing at the graduate level where English Proficient males earn over $70,000, nearly 50% more than their female counterparts.
