---
title: "Insights"
format: html
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
```

```{r}
# Read data
surveys <- read_rds(file = "data/usa_2021_23.rds")
```

```{r}
# Income by education and gender
surveys |>
  ggplot(aes(x = educ_level, y = exp(log_income), fill = gender)) +  # Transform back
  geom_boxplot() +
  facet_wrap(~ lep, 
             labeller = labeller(lep = c(`1` = "Limited English", `0` = "English Proficient"))) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(face = "bold")) +
  labs(
    title = "Income Distribution by Education Level, LEP Status, and Gender",
    x = "Education Level",
    y = "Income (USD)",
    fill = "Gender"
  )
```

This income distribution plot demonstrates clear education and gender premiums, with graduate degree holders earning substantially more than those with less education and males consistently out-earning females across all groups. Limited English Proficiency appears to have a modest negative impact on income, particularly at higher education levels, while significant income inequality is evident through extensive high-income outliers reaching $750,000+ across all subgroups.

```{r}
# Education distribution by LEP status
surveys |>
  count(lep, educ_level) |>
  group_by(lep) |>
  mutate(prop = n / sum(n) * 100) |>
  ggplot(aes(x = educ_level, y = prop, fill = factor(lep))) +  # Add factor() here
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_fill_discrete(name = "LEP Status", 
                     labels = c("English Proficient", "Limited English")) +
  labs(y = "Percentage (%)",
       x = "Education Level",
       title = "Education Distribution by LEP Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This education distribution plot reveals a stark educational divide between LEP groups, with Limited English Proficiency individuals heavily concentrated in lower education levels (nearly 50% have less than high school education) while English proficient individuals are more evenly distributed across education levels and comprise the majority of those with graduate degrees (around 40%).
